\section{What is \textit{Da-Sein}?}
Dasein is existence, and the awareness of existence, present only in humans.

Dasein has a few major components. One of the major components of dasein is
knowing that it exists. Fish, as opposed to humans, are not aware of the
temporal nature of their lives. They swim, eat, and eventually die, but are not
consciously thinking about the fact that one day, they will die. Humans, on the
other hand, live their daily routine knowing that one day, they will die.

It is this knowledge of the temporal nature of our existence that brings about
the idea of care. We know that the most important time is death, and in the time
before you die you can do things. Life, in essence, has a "shot-clock" in which
you can either do things, or die.

Humans are born with the ability to care. Care is the reason that people do
things. The synthesis of care and understanding, with a constraint on time, is
what makes human's identity.

\begin{description}
  \item[Care] The want to do things, the reasoning behind them.
  \item[Understanding] Knowing that we will die.
  \item[Temporal Nature of Existence] There is a finite amount of time that we
    have to do things.
  \item[Identity] The synthesis of all of the above.
\end{description}

The goal, in making one's Dasein is to be authentic. We are inextricably with
others, (this is "being with others", something that all humans have), but we
strive to be different. When we are not authentic, we listen to the "they," but
when we are, we are listening to ourselves. There is use in consulting the
"they" because they can provide you with information, but you must make your own
decision.
