\chapter{Exploration}
\section{Introduction}
Major European powers were motivated to expand their land for social powers.
The two leading powers in Europe at this time were Spain and Portugal, but they
each had differing strategies:

\begin{description}
  \item[Spain] trade routes, venture funding, etc.
    \begin{itemize}
      \item Columbus convinved the monarchs of Spain that since he believed that
        the Earth was smaller than commonly held at the time.  He recieved
        funding and sailed, eventually hitting land.  He thought that he had
        reached the East Indies, naming the populants of North America, "The
        Indians".
    \end{itemize}
  \item[Portugal] trade routes (primarily), they sought to bypass trade
    merchants in Italy thereby bringing them a greater profit.
\end{description}

Eventually, a treaty known as the \textit{Treaty of Tordesilias} was reached,
which devided the claims of the new land.  Portugal got Brazil, whereas Spain
got the rest of North America.

\section{Disease and Death}
Because a lot of the people of the New World did not have the same immunities as
that of those in the Old World, many were struck with disease unknown to them
and as a result, died.

\section{Slavery}
Many of the Spanish and Portugues colonists introduced a system of slavery.
This heavily influenced the system of Triangular Trade in the Atlantic.

\section{The Columbian Exchange}
The settlers of the New World acted as carriers for several different elements
of culture moving in both directions.  From the Old World, the New World
received horses, goats, cows, chickens, coffee, and more.  In addition, many
diseases were transported back and forth, which had a lasting impact on
diminishing the population in the New World.

\section{Spanish Exploitation}
The settlers from Spain were highly interested in Silver.  As such, they
developed a system known as the \textit{encomienda} that would exploit the labor
of native people to extract gold and silver and ship it to spain.  Spanish
settlers were granted tracts of land and the right to extract labor from local
residents.  This system closesly resembles Old World feudalism.

Eventually, this system was replaced with the \textit{repartimiento}.

\section{Motivation for Conquest}
\begin{description}
  \item[The Crusades] Since the Crusades are considered the death of the
    European feudal system, they were in dire need of foreign trade goods.  They
    wanted to circumvent the Italian city-states and make a higher profit off of
    their traded goods.
  \item[The Black Death] Similiarly to the Crusades, the Black Death also played
    a role in weakining the feudal system of the Old World.
  \item[The Renaissance] Brought upon an exploratory sense in the population of
    Europe.
  \item[The Protestant and Catholic reformations] renewed sense for members of
    society to spread their religion to new practicioners.
  \item[Technological Advances] the printing press helped spread information,
    whereas the compass, astrolabe, and quadrant all helped people to navigate.
\end{description}

\section{The Impact on Europe}
The influx of precious metals dramatically changed the European economy, and
helped facilitate the transition from feudalism to capitalism.

\subsection{The Columbian Exchange}
New crop and livestock were brought into Europe, such as turkey, corn, potato,
cacao, etc.  Spanish sailors also brought syphilis back to Europe, a result of
their sexual encounters with women of the New World.  Tabacco was also
introduced.

\subsection{Economic Impact}
Though one would think otherwise, the fruits of exploration did not impact Spain
in a positive way.  The influx of gold and silver caused hyper inflation during
the 1500s, and made ordinary items much more expensive.  As a result, taxes were
raised more than 5x in 1500 as well.  Spain went into debt to pay the military
expenditures that it had incurred from fighting in the new world.

\section{The Joint-Stock Company}
A Joint-Stock company is a form of a limited liability company where
shareholders own a proportinate amount of the compnay, directly varying with the
amount of stock that they own.  They are only responsible for the part that they
own.

\section{White Superiority}
A long-standing set of racist ideals helped settlers exploit non-whites for
labor.  This hirearchy had a variety of influences.
